---
title: "Medical Insurance Cost Prediction"
subtitle: "Supervised Learning - Assignment 1"
author: "Cesaire Tobias"
date: "April 29, 2025"
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
    lof: true
    lot: true
    documentclass: article
    geometry: margin=1in
    fontsize: 11pt
    linestretch: 1.2
execute:
  echo: false
  warning: false
  message: false
---

```{r setup, include=FALSE}
# Load required libraries - although we use the namespace convention the below indicates the required packages to run this file

library(knitr)       # For dynamic report generation
library(dplyr)       # For data manipulation
library(tidyr)       # For data tidying
library(ggplot2)     # For data visualisation
library(gridExtra)   # For arranging multiple plots
library(kableExtra)  # For enhanced table formatting
library(scales)      # For formatting plot scales
library(corrplot)    # For correlation visualisation
library(MASS)        # For statistical functions like stepAIC
library(glmnet)      # For regularised regression models

# Set default chunk options for the entire document
knitr::opts_chunk$set(
  echo = FALSE,      # Don't show code in final output
  warning = FALSE,   # Suppress warnings
  message = FALSE,   # Suppress messages
  fig.align = "center",  # Center figures
  fig.pos = "H",     # Position figures exactly here
  out.width = "80%", # Set figure width
  results = "asis"   # Output results as-is
)
```

```{r load-data}
rm(list = ls())
base_data_dir <- "./inst/extdata"

github_repo <- "https://github.com/ces0491/sl-assignment1.git"
github_data_url <- "https://github.com/ces0491/sl-assignment1/raw/refs/heads/main/data/insurance_data.RData"
kaggle_url <- 'kagglehub.dataset_download("mirichoi0218/insurance")'

my_seed <- 756
sample_size <- 500

#-------------------------------------------------------------------------

# read in full data either from file or github repo
if (file.exists(file.path(base_data_dir, "insurance.csv"))) {
  fulldata <- read.csv(file.path(base_data_dir, "insurance.csv"))
} else {
  con <- url(github_data_url)
  load(con)
  close(con)
}

# sample the data
set.seed(my_seed)
insurance_data <- fulldata[sample(nrow(fulldata), sample_size), ]

# save the sampled data
if (dir.exists(base_data_dir)) {
  write.csv(insurance_data, file.path(base_data_dir, "my_insurance.csv"), row.names = FALSE)
}
```

# Introduction

This report applies linear regression techniques to predict medical insurance costs (charges) in the United States (US) healthcare system. We seek to construct a linear supervised learning model to identify key determinants of individual medical costs billed by health insurance companies for patients in the US. In doing so, we propose some solutions to address the following questions for stakeholders:

-   **Patients**: Which personal factors most impact insurance charges?
-   **Insurers**: How can risk be accurately quantified for premiums?
-   **Policymakers**: Which public health initiatives can reduce costs?

The dataset includes 500 US patients with the following 6 features:

-   **age**: Integer, primary beneficiary's age
-   **sex**: Factor, gender of insurance contractor (female/male)
-   **bmi**: Continuous, body mass index
-   **children**: Integer, number of dependents
-   **smoker**: Factor, smoking status (yes/no)
-   **region**: Factor, US residential area (northeast, southeast, southwest, northwest)

The target variable is **charges** (medical costs billed). We aim to build linear models that predict charges based on the aforementioned characteristics, identifying key factors and interactions.

# Data Import and Preparation

We use a sub-sample of 500 observations, randomly selected with seed `r my_seed` for reproducibility.  
The `.RData` file containing both the full insurance data set (`fulldata`) and sampled data (`my_data`) can be found [here](https://github.com/ces0491/sl-assignment1/raw/refs/heads/main/data/insurance_data.RData).  
The original dataset can be found on Kaggle, [here](https://www.kaggle.com/datasets/mirichoi0218/insurance).  
Please see the GitHub repo [here](https://github.com/ces0491/sl-assignment1.git) for the full codebase.

```{r convert-factors}
if (!exists("insurance_data") || any(is.na(insurance_data))) {
  stop("Insurance data is missing or contains NAs")
}

# Convert character columns to factors and ensure no. of children is integer
insurance_data <- dplyr::mutate(
  insurance_data,
  sex = as.factor(sex),
  smoker = as.factor(smoker),
  region = as.factor(region),
  children = as.integer(children)
)
```

# Exploratory Data Analysis

## Data Overview

A sample of the dataset is shown below:

```{r show-data-head}
# Display top 5 rows of data
top_n_sample <- insurance_data %>%
  dplyr::top_n(5) %>%
  dplyr::mutate(dplyr::across(c(age, bmi, children, charges), ~round(.x, 2)))

# Create formatted table
kableExtra::kbl(
  top_n_sample,
  caption = "Insurance Dataset Sample (First 5 Rows)",
  format = "latex",
  booktabs = TRUE,
  align = "c",
  digits = 2
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 9,
    position = "center"
  ) %>%
  knitr::asis_output()
```

Summary statistics for numerical variables:

```{r summary-stats}
# Calculate summary statistics for numerical variables
summ_stats <- insurance_data %>%
  dplyr::select(age, bmi, children, charges) %>% # Select numeric columns
  dplyr::summarise(
    dplyr::across(
      .cols = dplyr::everything(), # Apply to all columns
      .fns = list(                 # Calculate multiple summ stats
        Min = ~min(.x, na.rm = TRUE),
        Q1 = ~quantile(.x, 0.25, na.rm = TRUE),
        Median = ~median(.x, na.rm = TRUE),
        Mean = ~mean(.x, na.rm = TRUE),
        Q3 = ~quantile(.x, 0.75, na.rm = TRUE),
        Max = ~max(.x, na.rm = TRUE)
      ),
      .names = "{.col}_{.fn}" # Name columns as variable_statistic
    )
  ) %>% # Reshape data for better presentation
  tidyr::pivot_longer(
    cols = dplyr::everything(),
    names_to = c("Variable", "Statistic"),
    names_sep = "_",
    values_to = "Value"
  ) %>%
  tidyr::pivot_wider(names_from = Statistic, values_from = Value) %>%
  dplyr::mutate(
    Variable = dplyr::recode(  # Rename variables for presentation
      Variable,
      "age" = "Age",
      "bmi" = "BMI",
      "children" = "Children",
      "charges" = "Charges"
    )
  )

# Create formatted table
kableExtra::kbl(
  summ_stats,
  caption = "Summary Statistics of Numerical Variables",
  format = "latex",
  booktabs = TRUE,
  align = "c",
  digits = 2
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 9,
    position = "center"
  ) %>%
  knitr::asis_output()
```

Categorical variable distributions:

```{r categorical-counts}
sex_table <- table(insurance_data$sex)
smoker_table <- table(insurance_data$smoker)
region_table <- table(insurance_data$region)

cat_summary <- data.frame(
  Variable = c(
    "Sex (female)", "Sex (male)",
    "Smoker (no)", "Smoker (yes)",
    "Region (northeast)", "Region (northwest)",
    "Region (southeast)", "Region (southwest)"
  ),
  Count = c(
    sex_table["female"], sex_table["male"],
    smoker_table["no"], smoker_table["yes"],
    region_table["northeast"], region_table["northwest"],
    region_table["southeast"], region_table["southwest"]
  ),
  Percentage = c(
    sex_table["female"]/sum(sex_table)*100, sex_table["male"]/sum(sex_table)*100,
    smoker_table["no"]/sum(smoker_table)*100, smoker_table["yes"]/sum(smoker_table)*100,
    region_table["northeast"]/sum(region_table)*100, region_table["northwest"]/sum(region_table)*100,
    region_table["southeast"]/sum(region_table)*100, region_table["southwest"]/sum(region_table)*100
  )
) %>%
  stats::na.omit()

kableExtra::kbl(
  cat_summary,
  caption = "Summary of Categorical Variables",
  format = "latex",
  booktabs = TRUE,
  align = "c",
  digits = 1
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 9,
    position = "center"
  ) %>%
  knitr::asis_output()
```

Key characteristics:

-   **Age**: 18–65 years, median `r median(insurance_data$age)` years
-   **Sex**: `r round(sex_table["female"]/sum(sex_table)*100, 1)`% female
-   **BMI**: Median `r round(median(insurance_data$bmi), 1)` (overweight range: 25–30)
-   **Smoker**: `r round(smoker_table["yes"]/sum(smoker_table)*100, 1)`% smokers
-   **Region**: Evenly distributed
-   **Children**: Mostly 0–1, max `r max(insurance_data$children)`

-   **Charges**: \$`r scales::comma(min(insurance_data$charges))` to \$`r scales::comma(max(insurance_data$charges))` (right-skewed)

## Visual Analysis

```{r charges-distribution, fig.cap="Distribution of Insurance Charges", fig.width=5, fig.height=3}
ggplot2::ggplot(insurance_data, ggplot2::aes(x = charges)) +
  ggplot2::geom_histogram(bins = 30, fill = "steelblue", color = "white") +
  ggplot2::theme_minimal() +
  ggplot2::labs(x = "Charges ($)", y = "Frequency") +
  ggplot2::scale_x_continuous(labels = scales::comma)
```

Charges are right-skewed, with most values between \$5,000–\$15,000, and some exceeding \$40,000.

```{r fig-interaction-viz, fig.cap="Charges by Smoking Status and BMI vs. Charges", fig.width=7, fig.height=3}
p1 <- ggplot2::ggplot(insurance_data, ggplot2::aes(x = smoker, y = charges, fill = smoker)) +
  ggplot2::geom_boxplot() +
  ggplot2::theme_minimal() +
  ggplot2::labs(x = "Smoker", y = "Charges ($)") +
  ggplot2::scale_y_continuous(labels = scales::comma) +
  ggplot2::scale_fill_brewer(palette = "Set1") +
  ggplot2::theme(legend.position = "none")

p2 <- ggplot2::ggplot(insurance_data, ggplot2::aes(x = bmi, y = charges, color = smoker)) +
  ggplot2::geom_point(alpha = 0.5) +
  ggplot2::geom_smooth(method = "lm", se = FALSE) + # Add linear trend line
  ggplot2::theme_minimal() +
  ggplot2::labs(x = "BMI", y = "Charges ($)") +
  ggplot2::scale_y_continuous(labels = scales::comma) +
  ggplot2::scale_color_brewer(palette = "Set1")

gridExtra::grid.arrange(p1, p2, ncol = 2)
```

Smokers face 3–4 times higher charges. BMI has a stronger effect on charges for smokers, indicating an interaction.

```{r age-charges, fig.cap="Age vs. Charges", fig.width=5, fig.height=3}
# Create scatterplot of age vs charges with smoothing line
ggplot2::ggplot(insurance_data, ggplot2::aes(x = age, y = charges, color = smoker)) +
  ggplot2::geom_point(alpha = 0.5) +  # Semi-transparent points
  ggplot2::geom_smooth(method = "lm", se = TRUE, color = "black") +  # Overall trend line
  ggplot2::theme_minimal() +
  ggplot2::labs(x = "Age (years)", y = "Charges ($)") +
  ggplot2::scale_y_continuous(labels = scales::comma) +  # Format y-axis
  ggplot2::scale_color_brewer(palette = "Set1")  # Color by smoking status
```

Age has a positive linear relationship with charges.

```{r children-charges, fig.cap="Children vs. Charges", fig.width=5, fig.height=3}
ggplot2::ggplot(insurance_data, ggplot2::aes(x = as.factor(children), y = charges)) +
  ggplot2::geom_boxplot(fill = "lightblue") +
  ggplot2::theme_minimal() +
  ggplot2::labs(x = "Number of Children", y = "Charges ($)") +
  ggplot2::scale_y_continuous(labels = scales::comma)
```

More children slightly increase charges, less impactful than age or smoking.

```{r correlation-matrix, fig.cap="Correlation Matrix", fig.width=4, fig.height=4}
correlation_matrix <- cor(insurance_data[, c("age", "bmi", "children", "charges")])
corrplot::corrplot(
  correlation_matrix,
  method = "circle",
  type = "upper",
  tl.col = "black",
  tl.srt = 45,
  addCoef.col = "black",
  col = colorRampPalette(c("#6D9EC1", "white", "#E46726"))(200),
  mar = c(0, 0, 2, 0),
  title = "Correlation Matrix of Numerical Variables"
)
```

Correlations: age (r = `r round(correlation_matrix['age','charges'], 2)`), BMI (r = `r round(correlation_matrix['bmi','charges'], 2)`), children (r = `r round(correlation_matrix['children','charges'], 2)`) with charges; low multicollinearity.

## Key Findings

-   **Smoking**: Dominant factor, significantly increasing charges
-   **Age**: Linear increase in charges
-   **BMI**: Stronger effect for smokers
-   **Children, Region and Sex**: Modest effect
-   **Charges**: Right-skewed distribution

# Model Building

## Data Splitting

```{r data-split}
set.seed(my_seed) # Set seed for reproducible random sampling
train_split <- 0.8 # Define train/test split ratio

# Create random indices for training data — sample without replacement
train_indices <- sample(nrow(insurance_data), train_split * nrow(insurance_data))

# Split data into training and testing sets
train_data <- insurance_data[train_indices, ]
test_data <- insurance_data[-train_indices, ]
```

| `r paste0("Training set: ", nrow(train_data), " observations (", scales::percent(train_split), ") of ", nrow(insurance_data), " in the total sample")`
| `r paste0("Testing set: ", nrow(test_data), " observations")`

## Full Regression Model

```{r full-model}
# Fit linear regression model with all predictors
full_model <- lm(charges ~ age + sex + bmi + children + smoker + region, data = train_data)
full_model_summary <- summary(full_model)
```

```{r full-model-table}
# Create coefficient table from model summary
coef_table <- data.frame(
  Term = rownames(full_model_summary$coefficients),
  Coefficient = round(full_model_summary$coefficients[,1], 2),
  Std_Error = round(full_model_summary$coefficients[,2], 2),
  t_value = round(full_model_summary$coefficients[,3], 2),
  p_value = round(full_model_summary$coefficients[,4], 3)
) %>%
  dplyr::mutate(
    Significant = ifelse(p_value < 0.05, "*", "")
  )

# Format neatly
kableExtra::kbl(
  coef_table,
  caption = "Full Regression Model Coefficients",
  format = "latex",
  booktabs = TRUE,
  align = "c",
  digits = c(0, 2, 2, 2, 3, 0),
  col.names = c("Term", "Coefficient", "Std. Error", "t value", "p value", ""),
  escape = FALSE
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 9,
    position = "center"
  ) %>%
  kableExtra::column_spec(5, bold = coef_table$p_value < 0.05) %>%
  knitr::asis_output()
```

R² = `r round(full_model_summary$r.squared, 3)`. Smoking, age, and BMI are significant predictors.

```{r full-model-predict}
# Make predictions on test data
full_model_pred <- predict(full_model, newdata = test_data)

# Calculate performance metrics
full_model_mse <- mean((test_data$charges - full_model_pred)^2)  # Mean squared error
full_model_rmse <- sqrt(full_model_mse)  # Root mean squared error
full_model_mae <- mean(abs(test_data$charges - full_model_pred))  # Mean absolute error
full_model_r2 <- 1 - sum((test_data$charges - full_model_pred)^2) / sum((test_data$charges - mean(test_data$charges))^2)  # R-squared
```

Test metrics: RMSE \$`r scales::comma(full_model_rmse)`, MAE \$`r scales::comma(full_model_mae)`, R² `r round(full_model_r2, 3)`.

# Model Improvement

## Variable Selection and Regularization

We explored stepwise (AIC/BIC), Ridge, and LASSO regression.

```{r model-improvement}
# Stepwise model selection using AIC
step_model_aic <- MASS::stepAIC(full_model, direction = "backward", trace = FALSE)

# Stepwise model selection using BIC (more stringent penalty -> simpler models)
step_model_bic <- MASS::stepAIC(full_model, direction = "backward", k = log(nrow(train_data)), trace = FALSE)

# Calculate metrics for AIC model
step_aic_pred <- predict(step_model_aic, newdata = test_data)
step_aic_mse <- mean((test_data$charges - step_aic_pred)^2)
step_aic_rmse <- sqrt(step_aic_mse)
step_aic_r2 <- 1 - sum((test_data$charges - step_aic_pred)^2) / 
  sum((test_data$charges - mean(test_data$charges))^2)

# Calculate metrics for BIC model
step_bic_pred <- predict(step_model_bic, newdata = test_data)
step_bic_mse <- mean((test_data$charges - step_bic_pred)^2)
step_bic_rmse <- sqrt(step_bic_mse)
step_bic_r2 <- 1 - sum((test_data$charges - step_bic_pred)^2) / sum((test_data$charges - mean(test_data$charges))^2)

# Prepare data for regularized regression
##  Convert categorical variables to dummy variables and remove intercept (first column)
x_train <- model.matrix(charges ~ age + sex + bmi + children + smoker + region, train_data)[,-1] # Regularization methods require numeric matrices as input. model.matrix() automatically handles factor variables.
y_train <- train_data$charges
x_test <- model.matrix(charges ~ age + sex + bmi + children + smoker + region, test_data)[,-1]
y_test <- test_data$charges

# Fit Ridge model using cross-validation to choose lambda (regularization strength)
set.seed(my_seed)
cv_ridge <- glmnet::cv.glmnet(x_train, y_train, alpha = 0, nfolds = 10)  # alpha=0 for ridge
lambda_ridge <- cv_ridge$lambda.min  # Optimal lambda

# Refit the model with optimal lambda
ridge_model <- glmnet::glmnet(x_train, y_train, alpha = 0, lambda = lambda_ridge)
ridge_pred <- predict(ridge_model, s = lambda_ridge, newx = x_test)

# Evaluate Ridge model performance
ridge_mse <- mean((y_test - ridge_pred)^2)
ridge_rmse <- sqrt(ridge_mse)
ridge_r2 <- 1 - sum((y_test - ridge_pred)^2) / sum((y_test - mean(y_test))^2)

# Ridge regression shrinks coefficients toward zero but keeps all variables in the model. This helps with multicollinearity and overfitting.

# LASSO regression with cross-validation
cv_lasso <- glmnet::cv.glmnet(x_train, y_train, alpha = 1, nfolds = 10)  # alpha=1 for LASSO
lambda_lasso <- cv_lasso$lambda.min  # Optimal lambda
lasso_model <- glmnet::glmnet(x_train, y_train, alpha = 1, lambda = lambda_lasso)
lasso_pred <- predict(lasso_model, s = lambda_lasso, newx = x_test)

# Evaluate LASSO model performance
lasso_mse <- mean((y_test - lasso_pred)^2)
lasso_rmse <- sqrt(lasso_mse)
lasso_r2 <- 1 - sum((y_test - lasso_pred)^2) / sum((y_test - mean(y_test))^2)
```

```{r stepwise-features}
# Extract features selected by each method
step_aic_features <- names(coef(step_model_aic))
step_bic_features <- names(coef(step_model_bic))
lasso_coefs <- coef(lasso_model, s = lambda_lasso)
lasso_features <- rownames(lasso_coefs)[lasso_coefs[,1] != 0]  # Non-zero coefficients

# Since ridge regression shrinks coefficients toward zero but never actually sets them exactly to zero, all predictors remain in the model and we exclude from comparing with feature selection methods.

# Create comparison table of selected features
feature_comparison <- data.frame(
  Feature = c("Intercept", "Age", "Sex (male)", "BMI", "Children", "Smoker (yes)",
              "Region (northwest)", "Region (southeast)", "Region (southwest)"),
  Full_Model = rep("x", 9),  # All features included in full model
  Stepwise_AIC = c(
    ifelse("(Intercept)" %in% step_aic_features, "x", ""),
    ifelse("age" %in% step_aic_features, "x", ""),
    ifelse("sexmale" %in% step_aic_features, "x", ""),
    ifelse("bmi" %in% step_aic_features, "x", ""),
    ifelse("children" %in% step_aic_features, "x", ""),
    ifelse("smokeryes" %in% step_aic_features, "x", ""),
    ifelse("regionnorthwest" %in% step_aic_features, "x", ""),
    ifelse("regionsoutheast" %in% step_aic_features, "x", ""),
    ifelse("regionsouthwest" %in% step_aic_features, "x", "")
  ),
  Stepwise_BIC = c(
    ifelse("(Intercept)" %in% step_bic_features, "x", ""),
    ifelse("age" %in% step_bic_features, "x", ""),
    ifelse("sexmale" %in% step_bic_features, "x", ""),
    ifelse("bmi" %in% step_bic_features, "x", ""),
    ifelse("children" %in% step_bic_features, "x", ""),
    ifelse("smokeryes" %in% step_bic_features, "x", ""),
    ifelse("regionnorthwest" %in% step_bic_features, "x", ""),
    ifelse("regionsoutheast" %in% step_bic_features, "x", ""),
    ifelse("regionsouthwest" %in% step_bic_features, "x", "")
  ),
  LASSO = c(
    ifelse("(Intercept)" %in% lasso_features, "x", ""),
    ifelse("age" %in% lasso_features, "x", ""),
    ifelse("sexmale" %in% lasso_features, "x", ""),
    ifelse("bmi" %in% lasso_features, "x", ""),
    ifelse("children" %in% lasso_features, "x", ""),
    ifelse("smokeryes" %in% lasso_features, "x", ""),
    ifelse("regionnorthwest" %in% lasso_features, "x", ""),
    ifelse("regionsoutheast" %in% lasso_features, "x", ""),
    ifelse("regionsouthwest" %in% lasso_features, "x", "")
  )
)

# Create formatted table
kableExtra::kbl(
  feature_comparison,
  caption = "Feature Selection Comparison",
  format = "latex",
  booktabs = TRUE,
  align = "c",
  col.names = c("Feature", "Full Model", "Stepwise (AIC)", "Stepwise (BIC)", "LASSO")
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 9,
    position = "center"
  ) %>%
  knitr::asis_output()
```

## Interaction Model

We include a BMI-smoking interaction term based on our findings in @fig-interaction-viz where we uncovered that BMI's effect on charges may not be constant across smoker vs. non-smoker groups.

```{r interaction-model}
interaction_model <- lm(charges ~ age + sex + bmi + children + smoker + region + bmi:smoker, data = train_data)
interaction_summary <- summary(interaction_model)

bmi_effect <- coef(interaction_model)["bmi"] # Effect of BMI for non-smokers
smoking_effect <- coef(interaction_model)["smokeryes"] # Effect of smoking
interaction_effect <- coef(interaction_model)["bmi:smokeryes"] # How BMI's effect changes for smokers
bmi_effect_smokers <- bmi_effect + interaction_effect # Total BMI effect for smokers

interaction_pred <- predict(interaction_model, newdata = test_data)
interaction_mse <- mean((test_data$charges - interaction_pred)^2)
interaction_rmse <- sqrt(interaction_mse) #avg prediction error
interaction_r2 <- 1 - sum((test_data$charges - interaction_pred)^2) / sum((test_data$charges - mean(test_data$charges))^2)
```

```{r interaction-model-summary}
int_coef_table <- data.frame(
  Term = rownames(interaction_summary$coefficients),
  Coefficient = round(interaction_summary$coefficients[,1], 2),
  Std_Error = round(interaction_summary$coefficients[,2], 2),
  t_value = round(interaction_summary$coefficients[,3], 2),
  p_value = round(interaction_summary$coefficients[,4], 3)
) %>%
  dplyr::mutate(
    Significant = ifelse(p_value < 0.05, "*", "")
  )

kableExtra::kbl(
  int_coef_table,
  caption = "Interaction Model Coefficients",
  format = "latex",
  booktabs = TRUE,
  align = "c",
  digits = c(0, 2, 2, 2, 3, 0),
  col.names = c("Term", "Coefficient", "Std. Error", "t value", "p value", ""),
  escape = FALSE
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 9,
    position = "center"
  ) %>%
  kableExtra::column_spec(5, bold = int_coef_table$p_value < 0.05) %>%
  knitr::asis_output()
```

The interaction is significant (p \< 0.001). BMI effect: \$`r round(bmi_effect, 0)` (non-smokers), \$`r round(bmi_effect_smokers, 0)` (smokers).
The different slopes shown below, illustrate the interaction - BMI’s effect on charges is steeper for smokers:
```{r interaction-visual, fig.cap="BMI-Smoking Interaction Effect", fig.width=5, fig.height=3}
new_data <- tidyr::expand_grid(
  bmi = seq(min(insurance_data$bmi), max(insurance_data$bmi), length.out = 100),
  smoker = c("yes", "no"),
  age = median(insurance_data$age),
  sex = "female",
  children = median(insurance_data$children),
  region = "southeast"
)

new_data$predicted_charges <- predict(interaction_model, newdata = new_data)

ggplot2::ggplot(new_data, ggplot2::aes(x = bmi, y = predicted_charges, color = smoker)) +
  ggplot2::geom_line(size = 1.2) +
  ggplot2::theme_minimal() +
  ggplot2::labs(
    title = "Predicted Charges by BMI and Smoking Status",
    subtitle = paste0(
      "BMI effect: $", round(bmi_effect), " (non-smokers), $",
      round(bmi_effect_smokers), " (smokers)"
    ),
    x = "BMI",
    y = "Predicted Charges ($)",
    color = "Smoker"
  ) +
  ggplot2::scale_y_continuous(labels = scales::comma) +
  ggplot2::scale_color_brewer(palette = "Set1") +
  ggplot2::annotate(
    "text",
    x = 30,
    y = 15000,
    label = paste0("Slope = $", round(bmi_effect)),
    color = "#E41A1C",
    fontface = "bold"
  ) +
  ggplot2::annotate(
    "text",
    x = 30,
    y = 45000,
    label = paste0("Slope = $", round(bmi_effect_smokers)),
    color = "#377EB8",
    fontface = "bold"
  )
```

## Model Comparison

```{r model-comparison}
model_comparison <- data.frame(
  Model = c("Full Model", "Stepwise (AIC)", "Stepwise (BIC)", "Ridge", "LASSO", "Interaction Model"),
  RMSE = c(full_model_rmse, step_aic_rmse, step_bic_rmse, ridge_rmse, lasso_rmse, interaction_rmse),
  R_Squared = c(full_model_r2, step_aic_r2, step_bic_r2, ridge_r2, lasso_r2, interaction_r2)
) %>%
  dplyr::arrange(RMSE) %>%
  dplyr::mutate(
    Bold = ifelse(RMSE == min(RMSE), "bold", "plain")
  )

kableExtra::kbl(
  model_comparison[, c("Model", "RMSE", "R_Squared")],
  caption = "Model Performance Comparison",
  format = "latex",
  booktabs = TRUE,
  align = "c",
  digits = c(0, 2, 3)
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 9,
    position = "center"
  ) %>%
  kableExtra::row_spec(
    which(model_comparison$RMSE == min(model_comparison$RMSE)),
    bold = TRUE
  ) %>%
  knitr::asis_output()
```

```{r rmse-visualization, fig.cap="Model Comparison by RMSE", fig.width=5, fig.height=3}
ggplot2::ggplot(model_comparison, ggplot2::aes(x = forcats::fct_reorder(Model, -RMSE), y = RMSE/1000, fill = Model)) +
  ggplot2::geom_bar(stat = "identity") +
  ggplot2::theme_minimal() +
  ggplot2::labs(
    title = "Model Comparison by RMSE",
    subtitle = "Lower RMSE indicates better performance",
    x = "Model",
    y = "RMSE (in thousands)"
  ) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1)) +
  ggplot2::scale_fill_brewer(palette = "Blues") +
  ggplot2::geom_text(ggplot2::aes(label = round(RMSE/1000, 2)), vjust = -0.5) +
  ggplot2::coord_cartesian(ylim = c(0, max(model_comparison$RMSE/1000) * 1.15))
```

The interaction model performs best, reducing RMSE by \$`r round(full_model_rmse - interaction_rmse, 2)`.

# Residual Diagnostics

```{r residual-diagnostics}
residuals <- residuals(interaction_model)
fitted_values <- fitted(interaction_model)
```

```{r resid-fitted, fig.cap="Residuals vs. Fitted Values", fig.width=4, fig.height=3}
data.frame(Fitted = fitted_values, Residuals = residuals) %>%
  ggplot2::ggplot(ggplot2::aes(x = Fitted, y = Residuals)) +
  ggplot2::geom_point(alpha = 0.5) +
  ggplot2::geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  ggplot2::geom_smooth(method = "loess", se = FALSE, color = "blue") +
  ggplot2::theme_minimal() +
  ggplot2::labs(
    title = "Residuals vs Fitted Values",
    subtitle = "Linearity and homoscedasticity check",
    x = "Fitted Values",
    y = "Residuals"
  ) +
  ggplot2::scale_x_continuous(labels = scales::comma)
```

```{r residual-normality, fig.cap="Normal Q-Q Plot", fig.width=4, fig.height=3}
data.frame(Residuals = residuals) %>%
  ggplot2::ggplot(ggplot2::aes(sample = Residuals)) +
  ggplot2::stat_qq() +
  ggplot2::stat_qq_line(color = "red") +
  ggplot2::theme_minimal() +
  ggplot2::labs(
    title = "Normal Q-Q Plot",
    subtitle = "Normality check",
    x = "Theoretical Quantiles",
    y = "Sample Quantiles"
  )
```

```{r scale-location, fig.cap="Scale-Location Plot", fig.width=4, fig.height=3}
data.frame(Fitted = fitted_values, Residuals = sqrt(abs(residuals))) %>%
  ggplot2::ggplot(ggplot2::aes(x = Fitted, y = Residuals)) +
  ggplot2::geom_point(alpha = 0.5) +
  ggplot2::geom_smooth(method = "loess", se = FALSE, color = "red") +
  ggplot2::theme_minimal() +
  ggplot2::labs(
    title = "Scale-Location Plot",
    subtitle = "Homoscedasticity check",
    x = "Fitted Values",
    y = "√|Standardized Residuals|"
  ) %>%
  ggplot2::scale_x_continuous(labels = scales::comma)
```

Diagnostics show acceptable linearity and normality, with some heteroscedasticity at higher values, suggesting potential for log transformation or robust methods such as robust standard errors, weighted least squares, quantile regression, or generalised least squares (GLS) to address heteroscedasticity when model residuals show increasing variance.

# Conclusion

## Key Findings

-   **Best Model**: Interaction model, RMSE \$`r scales::comma(model_comparison$RMSE[model_comparison$Model == "Interaction Model"])`, R² `r round(interaction_r2, 3)`
-   **Key Predictors**: Smoking (\$20,000 increase), BMI-smoking interaction (6× stronger for smokers), age, children.
-   **Cost Implications**: A 50-year-old smoker (BMI 30) pays \$`r scales::comma(round(predict(interaction_model, data.frame(age=50, sex="male", bmi=30, children=0, smoker="yes", region="southeast"))))` vs. \$`r scales::comma(round(predict(interaction_model, data.frame(age=50, sex="male", bmi=30, children=0, smoker="no", region="southeast"))))` as a non-smoker.

## Practical Implications

1. **Individuals**:
   - Quitting smoking offers the largest potential insurance savings.
   - Weight management is particularly crucial for smokers, as the financial impact of BMI is much stronger.
   - The combined effect of not smoking and maintaining healthy BMI could reduce annual costs by more than $25,000 for some individuals.
   - Age-related cost increases are inevitable but relatively modest compared to lifestyle factors.

2. **Insurance Providers**:
   - Risk assessment should incorporate the interaction between smoking and BMI.
    + Predictive models, in general, should consider interaction terms to accurately estimate costs.
   - More personalized premium structures could better reflect the compounding effects of multiple risk factors.
   - Prevention programs focused on smoking cessation for overweight individuals could yield substantial cost savings.

3. **Healthcare Policy**:
   - Public health initiatives targeting smoking cessation among overweight and obese individuals could yield disproportionately large healthcare savings.
   - Programs might benefit from focusing efforts on regions with higher baseline costs
   - Weight management programs specifically designed for smokers could deliver enhanced value.

## Limitations and Future Work

1. **Data Limitations**:
   - The dataset may not capture all relevant factors affecting insurance costs.
   - Additional variables like pre-existing conditions, medication usage, or healthcare utilization patterns would likely improve predictions.
   - Cross-sectional data limits our ability to analyze how changes in risk factors affect charges over time.

2. **Modeling Considerations**:
   - Log transformation of charges might better address heteroscedasticity.
   - Additional interactions (e.g, age×BMI) could potentially improve model performance.
   - More sophisticated machine learning approaches might capture complex nonlinear relationships.
   - Separate models for different regions or age groups might provide more tailored predictions.

3. **Future Research Directions**:
   - Longitudinal studies to assess how changes in risk factors affect insurance costs over time.
   - Analysis of cost-effectiveness of targeted interventions based on the interaction patterns identified.
   - Exploration of additional interaction effects among demographic and health variables.
   - Development of personalized risk calculators incorporating the identified factors.

Our analysis provides some insight into the key drivers of health insurance costs, highlighting the dramatic impact of smoking, its interaction with BMI, and the significant influence of age. These findings can inform more effective risk assessment practices, precision-targeted public health interventions, and improved insurance pricing models, ultimately benefiting both individuals and healthcare systems.